---
title: "Survey Exercises"
author: "Your Name"
date: "`r Sys.Date()`"
output: 
  html_document:
  toc: true
toc_float: true
number_sections: true
df_print: paged
---

## Getting Started

Ensure that you have `survey` installed on your machine

```{r setup, eval=FALSE}

install.packages(survey)

```


## Call The Libraries We Will Need

We'll use these libraries to help us explore the data

```{r}
library(survey) 
library(dplyr) 
```


Additionally, the `survey` package comes with some data examples. We will access them through the `data` function

```{r}
data(api)
```


We can now check our environment and see several new data sets, each representing different survey designs.


## Let's Investigate Our Data

WE can use `glimpse` from `dplyr` to help us explore our data

```{r}
glimpse(apisrs)
```

And of course the good old fashioned `print`

```{r}
print(apisrs)
```

## Specifying the Survey Design Object

The backbone of `survey` is the _survey.design_ object.

### Simple Random

```{r}
(svy_api_srs <- svydesign(ids = ~1, fpc = ~fpc, data = apisrs)) 
```

This special class of object helps our `svy` functions know how to calculate appropriate estimates (with the correct corrections for variances).

```{r}
class(svy_api_srs)
```

### Stratified Sample

```{r}
(svy_api_strat <- svydesign(ids= ~1, strata = ~stype, fpc = ~fpc, data = apistrat)) 
```


### Cluster Sample

```{r}
(svy_api_cluster <- svydesign(ids= ~dnum+snum, fpc = ~fpc1+fpc2, data = apiclus2))
```

## Performing Calculations with Survey Objects

All of the functions within `survey` being with the prefix `svy` and include the `survey.design` object as an argument. 

We can calculate means:

```{r}
svymean(~api00, svy_api_cluster)
```

And just for fun...

```{r}
cbind(mean(apiclus2[["api00"]]), sd(apiclus2[["api00"]]))
```

And quantiles

```{r}
svyquantile(x = ~api99+api00, svy_api_srs, quantile= c(0.25,.75))
```

Totals

```{r}
svytotal(x = ~stype, svy_api_srs)
```

Sometimes we want to do some subgroup analysis. We can use the `svyby` to do an analysis by another variable (like a `group_by` argument)

```{r}
(mean_score <- svyby( ~api99, ~stype, svymean, design = svy_api_cluster) )
```

From that, we can then use the `svycontrast` to create a constrast

```{r}
svycontrast(mean_score, quote(H/E))
```


Or add the new columns directly to the survey object with the `update` function:

```{r}
(svy_api_cluster <- update(svy_api_cluster, score_imp = api00/api99)) 
```

## Regression

```{r}
svyglm(score_imp~ meals + avg.ed, svy_api_cluster)
```

Additionally, we can perform subsetting using `subset` on the survey design object.

```{r}
svyglm(score_imp~ meals + avg.ed, 
       subset(svy_api_cluster, stype == "E")) 
```

## Post Survey Correction

### Post-stratification

First we need to make some fake data to simulate (really it is real data, we will just do some manipulations on it):

```{r}
df <- (MASS::survey) %>% 
  na.omit()

```

Let's check some of our outcomes

```{r}
prop.table(table(df$Sex)) 
```

Build our survey object

```{r}
survey_design_unweighted <- svydesign(ids = ~1, data =df)
```

We get a warning here, but thats ok. It basically indicates we didn't give the function any weight so everyone gets a weight of 1. 

But now we need to make some information about our population available to our analysis.

```{r}
(gender_dist <- data.frame( Sex = c("Female", "Male"), Freq = round(nrow(df) * c(.7, .3),0))) 
```

Now with this information about our population we can perform some post-stratification.


```{r}
(survey_design_weighted <- postStratify( survey_design_unweighted, 
                                         ~Sex, # Variable to Post-stratify
                                         gender_dist # Population Info
                                         )) 
```

And now we can check out results

```{r}
svymean(~Height, survey_design_unweighted)
svymean(~Height, survey_design_weighted)
```

Adding more variables:

```{r}
prop.table(table(df$W.Hnd))
```

More population information!

```{r}
(handed <- data.frame( W.Hnd = c("Left", "Right"), 
                       nrow(df) * c(.1, .9)))
```

### Apply the Rake

```{r}
survey_design_rake <- rake( survey_design_unweighted, 
                            sample.margins = list(~Sex, ~W.Hnd), 
                            population.margins = list(gender_dist,handed))

```

### Check the Weights

```{r}
summary(weights(survey_design_rake))
```

### Trim if Needed

There are all kinds of methods to look at weight trimming. Choose one to apply if necessary.

```{r}
median_wt <- median(weights(survey_design_rake))
IQR_wt <- IQR(weights(survey_design_rake))
```

```{r}
trimmed <-trimWeights(survey_design_rake, 
                      upper = median_wt + 2*IQR_wt, 
                      lower = median_wt - 2*IQR_wt)
```

## Communicating Results

Add the weights back to finalised data set!

```{r}
df_with_wts <- df %>% add_column(wts = weights(trimmed))
```

## Pre-view of `srvyr`

Provides some `tidyverse` syntax to `survey`. Read more at <https://cran.r-project.org/web/packages/srvyr/vignettes/srvyr-vs-survey.html>

```{r}
library(srvyr)

df_with_wts %>% 
  as_survey_design(ids = 1, wts = wts) %>% 
  group_by(W.Hnd) %>% 
  summarise(mu_height = survey_median(Height),
            mu_height_unwt = unweighted(median(Height)))
```

