---
title: "Advanced Data Analysis in R"
subtitle: "Survey Analsis in R"
author: "Michael DeWitt"
date: "2018-02-09 (Updated `r Sys.Date()`)"
output:
  beamer_presentation:
    keep_tex: false
    theme: metropolis
    slide_level: 2
    incremental: false
    includes:
      in_header: head.txt
fontsize: 12pt
classoption: compress
bibliography: my_bib.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = TRUE,
  cache = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)
library(tidyverse)
library(broom)
library(here)
set.seed(27)
options(
  tibble.print_max = 3,
  tibble.print_min = 3,
  width = 55
)
theme_set(theme_gray(base_size = 22))
```

# Survey Analysis in R

## What makes survey analysis different?

Survey analysis is _design based_

Often we talk about _probability or random samples_

These concepts make inferences really nice


## Properties of Design Based Surveys

A quick refresher ^[@lumley_complex_2010]  

>1. Every individual in the population must have a non-zero probability of ending up in the sample ($\pi_i$)  
>2. The probability of $\pi_i$ must be known for every individual in who does end up in the sample  
>3. Eevery pair of individuals in the sample must have a non-zero probability of both ending up in the sample ($\pi_{i,j}$ for the pair of individuals (i, j))  
>4. The probability $\pi_{i,j}$ must be known for every pair that does ebd up in the sample  



# Introducing the `survey` package

## A little about `survey`

[Thomas Lumley](https://www.stat.auckland.ac.nz/people/tlum005) developed the `survey` package

Initially a port of STATA's `svy` functions following a similar syntax

Can perform typical types of design based analysis

- Simple Random  
- Stratified   
- Clusters  
- Multi-stage  
- Repeated Measures  

## A little about `survey`

Perform post-survey corrections  
- post-stratification  
- raking (iterative proportional fitting)  
- calibration  

And more...!

# Diving into the software...

## Describing your model

The primary argument in `survey` is the [`svydesign`](https://www.rdocumentation.org/packages/survey/versions/3.35-1/topics/svydesign) function

```{r svydesign, eval=FALSE}
library(survey)

svydesign(ids = to specify clusters (~1 otherwise),
          probs = Sampling Probabilities if available, 
          strata = Strata membership if available,
          fpc = Finite Population Values, 
          data = Your Data Frame, 
          nest = T/F if there is nesting within your strata, 
          weights = Sampling Weights if available,
          pps = Probability Proportional to Size)
```

## Quick Note On `survey`

Many of the functions in `survey` utilise R "formula notation"

Indicates the tilde "~" must be used (e.g. `~cluster`)

## But Let's Try An Example

Let's try an example with the `api` dataset that is part of the `survey` package

This data set represents California Academic Performance Index

```{r}
library(survey)
library(dplyr)
data(api)
```

## Let's Inspect the Data

```{r}
head(apisrs)
```

## Specifying the Survey Object (SRS)

This is a simple random sample with finite population correct (since we know the population)

```{r}
(svy_api_srs <- svydesign(ids = ~1, 
                          fpc = ~fpc, 
                          data = apisrs))
```


## Trying With A Different Survey Design (Stratified)

In this case we have a stratified random sample (different school types)

```{r}
(svy_api_strat <- svydesign(ids= ~1, 
                            strata = ~stype, 
                            fpc = ~fpc, 
                            data = apistrat))
```

## Trying With A Different Survey Design (Cluster)

Two stage cluster sampling 40 school disticts then five schools within each district

- Stage 1 district cluster with population `fpc1`  
- Stage 2 district cluster with population `fpc2`

```{r}
(svy_api_cluster <- svydesign(ids= ~dnum+snum, 
                              fpc = ~fpc1+fpc2, 
                              data = apiclus2))
```

# Analysis with svy objects

## Correct Estimates

`survey` applies correct calculations given the survey design

```{r}
svymean(~api00, svy_api_cluster)
```

vs

```{r}
cbind(mean(apiclus2[["api00"]]), 
      sd(apiclus2[["api00"]]))
```

## Survey Functions

Functions in the `survey` package begin with the `svy` prefix

Utilise the formula notation

```{r}
svyquantile(x = ~api99+api00, 
            svy_api_srs, 
            quantile= c(0.25,.75))
```

## Calculating Contrasts

You can add contrasts with `svycontrast`

Say I wanted to look at the ratio of my high school score to my elementary school score

```{r}
# Mean
mean_score <- svyby( ~api99, ~stype, svymean, 
                       design = svy_api_cluster)
# Contrast ratio use `quote`
svycontrast(mean_score, quote(H/E))
```

## Adding Contrasts to the data

Use the `update` function to add new calculated fields to your survey design object

```{r}
(svy_api_cluster <- update(svy_api_cluster, 
                          score_imp = api00/api99))
```

## Adding Contrasts to the data

Now we can easily perform our analysis

```{r}
svyby(~score_imp, ~stype, svymean, 
        design = svy_api_cluster)
```

## Performing Regressions

```{r}
svyglm(score_imp~ meals + avg.ed, svy_api_cluster)
```


# Post-survey corrections

```{r}
df <- (MASS::survey) %>% na.omit()
```

```{r}
prop.table(table(df$Sex))
```

```{r}
survey_design_unweighted <- svydesign(ids = ~1, data =df)
```


```{r}
(gender_dist <- data.frame(Sex = c("Female", "Male"), Freq = round(nrow(df) * c(.55, .45),0)))
```

```{r}
survey_design_weighted <- postStratify(survey_design_unweighted, ~Sex, gender_dist)
```

```{r}
svymean(~Height, survey_design_unweighted)
```

```{r}
svymean(~Height, survey_design_weighted)
```


```{r}
prop.table(table(df$W.Hnd))
```

```{r}
(handed <- data.frame(W.Hnd = c("Left", "Right"), nrow(df) * c(.1, .9)))
```

```{r}
survey_design_rake <- rake(survey_design_unweighted, sample.margins = list(~Sex, ~W.Hnd), population.margins = list(gender_dist,handed))
```

```{r}
summary(weights(survey_design_rake))
```

## References

<!--
- Postratification and Ranking
- Trimming Weights
- Cluster
-->